### 2018-09-28 paper.tex
Wip

@@

\section{Developing Reusable Models}

@@

- How to choose a model boundary?

- How to define operations

\subsection{Customizing Models}

\subsubsection{Model Inheritance}

\subsubsection{Configuration Mixins}

\subsubsection{Modifying Operations}

\subsubsection{Modifying Flag Values}

@@

@@

\subsection{Patching Strategies}

>>>>

\subsection{Run Directories}
\label{sec:run-dir}

A \emph{run directory} is a directory used for a run. During a run
operation, the run directory is the system current working
directory. Operations should write run artifacts within the current
working directory to ensure that they are associated with the run.

By default, Guild AI generates a new directory for each run and
creates links to required resource sources within that directory. Once
the directory is prepared, Guild AI starts the run operating system
process, which then may access sources within the current working
directory.

Users may optionally specify an alternative run directory for a
run. This is commonly used by model developers during the process of
implementing model operation.s

@@

\section{Developing Models}

As discussed above, published models---deployed from installed
packages or as projects---can be used by running model
\emph{operations}. Operations generate \emph{artifacts} that can be
used in subsquent operations or as final outputs in an application
development workflow.

In this section we discuss how such models are created using Guild AI.

@@

When developing a model from scratch, a developer may ask two
questions:

\begin{itemize}
\item What is the defining characteristic of the model?
\item What can users do with the model?
\end{itemize}

The answers to these questions will inform the decision of what to
name the model and its operations. These values can be stubbed out in
a Guild file to provide a scaffold for implementation.

In many cases, this is simply a matter of selecting a model name and
creating a \verb|train| operation. Listing~\ref{lst:pet-stub}
illustrates a straightforward pet classifier.

\begin{lstlisting}[
    caption=Sample Guild file for a pet classifier,
    label={lst:pet-stub}]
model: pets
description:
  Pet classifier using a CNN
operations:
  train:
    description: Train the classifier
    main: train
    flags:
      images:
        description:
          Directory containing the pet
          images to train
        required: yes
      epochs:
        description:
          Number of epochs to train
        default: 100
\end{lstlisting}

With even a simple stub, a developer can formalize the model and what
it does. This stage clarifies the model's user interface before the
functionality is implemented.

With a basic skeleton in place, the developer can implement model
operations. In the case of the pet classifier
(Listing~\ref{lst:pet-stub}) this means implementing training support
in a file named \verb|train.py| that resides in the project directory.

Operations can be developed and tested early on using Guild AI. Each
time \verb|guild run| is used to test an operation, Guild AI generates
a distinct, trackable run. Runs may be compared over time to help
answer questions or resolve issues.

Guild AI provides two additional mechanisms to help model developers
implement operations. Both effect the \emph{run directory}
(Section~\ref{sec:run-dir}) used by Guild AI.

The first is the use the \verb|--run-dir| option
to the \verb|guild run| command. This option may be used to specify an
alternative directory for the run. This is helpful when the developer
doesn't want to generate new runs for early stage opertion
development.

The second mechanism is the use of the \verb|--stage| option to the
\verb|guild run| command. This option may be used to stage a run in an
alternative directory without running the operation. This is helpful
when the developer wants to study the run directory layout or to run
low level commands or scripts during the development process.

The developer can continue to evolve the model operations as needed to
satisfy user requirements. This may include new operations or
additional operation flags, which give users more flexibility when
running an operation. The developer may similarly add new models to
the Guild file.

\subsection{Adapting Existing Model Code}

In many cases, a developer will already have working model code she
wants to incorporate into a formalized model definition. As with the
role of \emph{packger} (Section~\ref{sec:package-maintainer}), a
developer may even be required to work with upstream sources she
cannot easily change. Guild AI provides a number of features that
support the adaptation of existing model code into usable models.

The operation \verb|main| attribute specifies a Python main module to
run. In many cases, a developer can simply reference an existing
Python script to incorporate it into an operation. In cases where a
script doesn't already exist---for example, the model code is in a
Jupyter Notebook and not readily usable---the developer should create
a new script that contains the applicable model code and use it for an
operation.

Operation \emph{flags} specified in the form \verb|NAME=VALUE| as
arguments to the \verb|guild run| command are passed to the Python
main module as command line arguments in the form
\verb|--NAME VALUE|. For example, assuming an operation \verb|train|
with a main module \verb|train|, the command
\verb|guild run train epochs=100| would be translated to the Python
command \verb|python -m train --epochs 100|. In this way developers
can parameterize scripts via operation flags.\footnote{It is standard
  practice to use Python's argparse module to support named command
  line options following the getopt convention (ref) in a Python
  script.}

If a flag name doesn't correspond to the option name expected
(e.g. the train script expected the option \verb|num_epochs|) the
developer can specify an alternative argument name for a flag using
the \verb|arg-name| attribute.

\subsection{Patching}

In cases where a developer cannot modify upstream sources or chooses
not to (e.g. to simplify ongoing mainteance) he may choose to patch
model sources for use in a model. Guild AI supports two methods of
patching.

The first method is to run a post processing script on required model
sources. This can be specified using a \verb|post-process| command for
a model source that applies patch files to upstream sources. Guild AI
will use the patched versions for related operations.

The second method is to wrap upstream scripts and monkey patch (ref)
the Python environment as needed to change behavior. An example of
this can be found in the \verb|gpkg.slim| package, which patches the
TensorFlow \verb|slim.losses.softmax_cross_entropy| module to add
support for balanced weight training. This technique successfully adds
a useful feature while avoiding the need to modify the upstream train
support in TF Slim.\footnote{While one could argue that it would be
  better to submit a pull request with the new functionality to
  upstream authors, there are always cases where this is not an
  option. Patching in these cases may be the only way to fix issues or
  add features.}

\subsection{Model Definition Reuse}

Guild AI provides a number of features that support model definition
reuse.

\subsubsection{Configuration Inheritance}

Guild files may contain \verb|config| section that contain arbitrary
configuration, which can be inherited by
models. Listing~\ref{lst:inheritance} defines two models, each
inheriting the properties of shared config. The shared config
\verb|model-base| defines the common model operations and uses
\emph{parameters} to define a model name and default learning rate.

\begin{lstlisting}[
    caption=Configuration inheritance,
    label={lst:inheritance}]
- config: model-base
  operations:
    train:
      main: train --model {{model-name}}
      flags:
        epochs: 100
        learning-rate: {{default-lr}}

- model: logreg
  extends: model-base
  params:
    model-name: logreg
    default-lr: 0.01

- model: cnn
  extends: model-base
  params:
    model-name: cnn
    default-lr: 0.0001
\end{lstlisting}

Configuation inheritance can be applied across Guild files---e.g. to
configuration defined in packages or other projects. This is useful
for extending model definitions from other
developers. Listing~\ref{lst:extend-slim} demonstrates how a model can
inherit all of the operations of another model with a single line of
configuration.

\begin{lstlisting}[
    caption=Extending PNASNet Mobile,
    label={lst:extend-slim}]
- model: my-classifier
  extends: gpkg.slim.models/pnasnet-mobile
\end{lstlisting}

Configuration inheritance has proven effective at managing complex
model configuration. This is evident in the \verb|gpkg.slim| and
\verb|gpkg.object-detection| packages, which make extensive use of
configuration inheritance to simplify their respective Guild files.

\subsubsection{Configuration Includes}

Configuration may also be reused by including shared configuration
files using \verb|include| objects. Listing~\ref{lst:shared} and
Listing~\ref{lst:include-shared} illustrate how includes can be used.

\begin{lstlisting}[
    caption=Shared configuration (e.g. defined in shared.yml),
    label={lst:shared}]
- config: model-base
  operations:
    train: train
\end{lstlisting}

\begin{lstlisting}[
    caption=Use of includes (e.g. defined in guild.yml),
    label={lst:include-shared}]
- include: shared.yml

- model: model-a
  extends: model-base

- model: model-b
  extends: model-base
\end{lstlisting}

Includes may also be used to merge shared configuration into other
configuration. Listing~\ref{lst:include-flags} illustrates how shared
configuration can be included into operation flags. The special
attribute \verb|$include| signifies that the corresponding value
defined in the referenced config should be merged into the current
value. The value for \verb|$include| can be a list to merge from
multiple sources. In the example, the value for \verb|learning-rate|
is redefined in the operation.

\begin{lstlisting}[
    caption=Including shared flag configuration,
    label={lst:include-flags}]
- config: default-train-flags
  flags:
    epochs: 100
    batch-size: 32
    learning-rate: 0.01

- model: sample
  operations:
    train:
      main: train
      flags:
        $include: default-train-flags
        learning-rate: 0.001
\end{lstlisting}

@@

The final method of model definition reuse described is that of
\emph{Python main module reuse}. Main modules---i.e. the modules
referenced in an opertion \verb|main| specification---may reside in
any installed Python package. To reuse the main module, a developer
need only include the package reference in the main specification.

Listing~\ref{lst:tflite_convert} illustates the use of the TensorFlow
module \verb|tflite_convert| to generte a TF Lite file for an
operation.

\begin{lstlisting}[
    caption=Reusing a Python main module,
    label={lst:tflite_convert}]
model: sample
operations:
  tflite:
    main: tensorflow/contrib/lite/python/tflite_convert
\end{lstlisting}

Developers may create packages of reusable modules that can be easily
used to implement model operations. For example, consider an
organization that runs a specialized inference server that hosts
models deployed using an HTTP POST interface. The model development
team might create a package \verb|server-utils| that contains a Python
main module \verb|server_utils.deploy| that performs the
deployment.

Listing~\ref{lst:deploy} illustrates how such a facility could be
easily incorporated into a model to enable deployments. In the
example, the model is defined in a package that requires the
\verb|server-utils| package, which will be automatically installed by
the command \verb|guild install classifiers|.

\begin{lstlisting}[
    caption=Quickly enabling model deploy,
    label={lst:deploy}]
- package: classifiers
  requires: [server-utils]

- model: resnet-classifier
  operations:
    train: resnet_train
    deploy: server_utils.deploy
    requires: trained-model
  resources:
    trained-model:
      sources:
        - operation: train
\end{lstlisting}

@@
